---
title: "Introduction to socratr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to socratr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# socratr

`socratr` reads and writes datasets on [Socrata](https://dev.socrata.com)
open data portals. It supports SODA 3 JSON and SODA 2 CSV downloads,
optional schema-based type coercion, and parallel fetches for large
datasets.

```{r setup}
library(socratr)
```

## Reading data

```{r read-json}
df <- read_socrata(
  "https://soda.demo.socrata.com/resource/4334-bgaj",
  app_token = Sys.getenv("SOCRATA_APP_TOKEN")
)
```

Use CSV for large public downloads, and `coerce = TRUE` for typed columns:

```{r read-csv}
df <- read_socrata(
  "https://soda.demo.socrata.com/resource/4334-bgaj",
  format = "csv",
  coerce = TRUE,
  app_token = Sys.getenv("SOCRATA_APP_TOKEN")
)
```

Filter with SoQL (do not include `LIMIT` / `OFFSET` — pagination is
handled internally):

```{r read-soql}
df <- read_socrata(
  "https://soda.demo.socrata.com/resource/4334-bgaj",
  soql = "SELECT region, magnitude WHERE magnitude > 3",
  app_token = Sys.getenv("SOCRATA_APP_TOKEN")
)
```

## Parallel reads

```{r read-parallel}
df <- read_socrata_parallel(
  "https://soda.demo.socrata.com/resource/4334-bgaj",
  format = "csv",
  max_active = 5L,
  app_token = Sys.getenv("SOCRATA_APP_TOKEN")
)
```

## Metadata and coercion

```{r metadata}
meta <- get_metadata(
  "https://soda.demo.socrata.com/resource/4334-bgaj",
  app_token = Sys.getenv("SOCRATA_APP_TOKEN")
)
meta$columns

df <- read_socrata(
  "https://soda.demo.socrata.com/resource/4334-bgaj",
  app_token = Sys.getenv("SOCRATA_APP_TOKEN")
)
df <- coerce_socrata_types(df, meta)
```

## Listing datasets

```{r list}
ls_socrata("soda.demo.socrata.com", limit = 10L)
```

## Writing data

Writing uses SODA 2. You need credentials with write access to the target
dataset.

```{r write}
write_socrata(
  dataframe    = data.frame(x = 1, y = 2),
  domain       = "soda.demo.socrata.com",
  dataset_id   = "xh6g-yugi",
  update_mode  = "UPSERT",
  socrata_user = Sys.getenv("SOCRATA_USER"),
  password     = Sys.getenv("SOCRATA_PASSWORD"),
  app_token    = Sys.getenv("SOCRATA_APP_TOKEN")
)
```

## Authentication

```{r auth}
Sys.setenv(
  SOCRATA_APP_TOKEN  = "your_app_token",
  SOCRATA_USER       = "your@email.com_or_api_key_id",
  SOCRATA_PASSWORD   = "your_password_or_api_secret"
)
```
